### 2015-02-17 doc/Milestone/milestone.tex
latest updates

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{Recognizing Strong Gravitational Lenses}

\author{Chris Davis\\
{\tt\small cpd@stanford.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Andrew McLeod\\
{\tt\small ajmcleod@stanford.edu}
}

\maketitle
%\thispagestyle{empty}

% %%%%%%%%% ABSTRACT
% \begin{abstract}
%    The ABSTRACT is to be in fully-justified italicized text, at the top
%    of the left-hand column, below the author and affiliation
%    information. Use the word ``Abstract'' as the title, in 12-point
%    Times, boldface type, centered relative to the column, initially
%    capitalized. The abstract is to be in 10-point, single-spaced type.
%    Leave two blank lines after the Abstract, then begin the main text.
%    Look at previous CVPR abstracts to get a feel for style and length.
% \end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

A consequence of Einstein's Theory of General Relativity is that mass bends the
path of light. When light passes through a particularly deep gravitational
potential, the deflections, called strong gravitational lenses, can produce
brilliant arcs and multiple images. The number and distribution of strong
lenses reflect the expansion history of the universe.

%-------------------------------------------------------------------------
\section{Problem Statement}

In this project we will use images collected by the Canada-France-Hawaii
Telescope Legacy Survey to analyze how Convolution Neural Networks can improve
automated detection of strong lens systems. From other graduate work (but not
coursework), we have the locations and categories of around one hundred and
twenty known strong lenses, three thousand large fields verified to contain no
strong lenses, six thousand simulated strong lenses, and several thousand
classifications by citizen-scientists of other potential strong lens systems.
These will form the core of our training and testing datasets; our metric will
be how well a CNN correctly identifies known and simulated lenses and non-lens
systems.

We would like to examine the following questions:
\begin{itemize}
\item{ Do we have enough data to reasonably train and test a CNN? Can we get
       around this by artificially inflating the data, e.g. by adding rotated
     images?}
\item{ What processing needs to be done on the data? What kind of scaling of
       pixel data is appropriate for automated detection? Should we compress
     the five different 'colors' to a smaller number of dimensions?}
   \item{ How do citizen-scientists do compared with this automated system?}
   \item{Can we use the results of citizen-scientists to train the CNN?}
\item{What sets of classifications are needed? Are we better served sticking
       to 'lens' and 'not lens', or should we use several classification
       categories ('lensed arcs', 'lensed multiple images', 'non-lens pixel
     noise')?}
\end{itemize}

%-------------------------------------------------------------------------
\section{Technical Approach}

From approximately 12000 fields of $440\times440\times3$ fields, we have
constructed ______ cutouts sized $96\times96\times3$. We

%-------------------------------------------------------------------------
\section{Intermediate/Preliminary Results}

% {\small
% \bibliographystyle{ieee}
% \bibliography{egbib}
% }

\end{document}

### 2015-02-17 doc/Milestone/milestone.tex
updating the milestone

@@
 A consequence of Einstein's Theory of General Relativity is that mass bends the
-path of light. When light passes through a particularly deep gravitational
-potential, the deflections, called strong gravitational lenses, can produce
-brilliant arcs and multiple images. The number and distribution of strong
-lenses reflect the expansion history of the universe.
+path of light. Most of the time the deflections are very small; the original
+'gravitational lens' that tested the veracity of Einstein's theory in the years
+after the first world war was the sun, which deflected light from stars behind
+it only a few seconds of arc. However, when light passes through a particularly
+deep gravitational potential (say, near the center of the dark matter halo of a
+galaxy cluster), the deflections can be particularly large, resulting in
+brilliant arcs and multiple images. These strong deflections due to light
+passing through a deep gravitational potential are termed strong gravitational
+lenses.
+
+The very existence of these potentials acts as a verification of the Theory of
+General Relativity, but they can also be used for much more. Strong
+gravitational lenses are one of the few ways to directly probe the distribution
+of dark matter, a particle (or possibly family of particles) that does not emit
+electromagnetic radiation but does have mass and hence interacts
+gravitationally with normal baryonic matter. This allows us to tally the mass
+of the largest gravitationally-bound structures in the universe, galaxy
+clusters, which can give us insight into the formation history of these massive
+objects. In this way, strong gravitational lenses can then tell us something
+about the expansion history of the universe, by setting limits on how massive
+the most massive objects in the universe can be. The properties of the bent
+light itself can also say much about that expansion history. When an object is
+strongly-lensed into multiple images, each image travels a different span of
+space and time. When an object does not vary much with time, these different
+path lengths have no practical import. However, if the object varies
+appreciably quickly (say it is a distant supermassive black hole at the center
+of a galaxy whose accretion disk emits high-energy radiation at varying rates)
+then these different path lengths can be used to pin down the rate of expansion
+of the universe.
+
+Unfortunately, for how useful strong gravitational lenses are, they are also
+extremely rare. A next generation optical survey like the Large Synoptic Survey
+Telescope can expect to find only ten thousand lenses in the whole sky, while
+it will find ten billion galaxies. Currently in astronomy there are only order
+hundreds of strong gravitational lenses known, mostly discovered by 'eyeball
+squads' of graduate students. The small number means that target criteria must
+be somewhat broad in order to maintain a relatively high completeness. Using
+reasonable target criteria to find strong lenses such as looking only at
+massive galaxies still means that nearly ten million objects will need to be
+inspected in the next generation in order to find those ten thousand lenses. A
+team of ten graduate students could expect to spend about 14 years looking at
+these objects. Computer algorithms are not much better: most current machine
+learning algorithms are woefully-underpowered for this task, and generally have
+poor completeness or poor purity -- and often poor both. Additionally, some
+algorithms are better at finding some types of lenses than others -- some
+perform well on the brilliant arcs, but poorly on the multiply-imaged objects,
+or vice versa. New algorithms need to be developed to find more strong
+gravitational lenses, and more strong lenses need to be found to power these
+algorithms.
+
+{\sc Space\,Warps}\xspace is a citizen-science initiative designed to overcome
+these two problems. The program has users examine images from the
+Canada-France-Hawaii Telescope Legacy Survey (CFHTLS) and vote on where they
+see lenses. Users are also assessed and trained with simulated lenses and known
+empty fields. By having thousands of users analyze a survey for short amounts
+of time each, it is hoped that a more complete sample of lenses can be
+discovered, which can then be fed into lens-finding algorithms to further
+improve their performance.
 
@@
 Telescope Legacy Survey to analyze how Convolution Neural Networks can improve
-automated detection of strong lens systems. From other graduate work (but not
-coursework), we have the locations and categories of around one hundred and
-twenty known strong lenses, three thousand large fields verified to contain no
-strong lenses, six thousand simulated strong lenses, and several thousand
-classifications by citizen-scientists of other potential strong lens systems.
-These will form the core of our training and testing datasets; our metric will
-be how well a CNN correctly identifies known and simulated lenses and non-lens
-systems.
+automated detection of strong lens systems. We will also assess the performance
+of citizen scientists by comparing our results to them. From other graduate
+work (but not coursework), we have the locations and categories of around one
+hundred and twenty known strong lenses, three thousand large fields verified to
+contain no strong lenses, six thousand simulated strong lenses, and several
+thousand classifications by citizen-scientists of other potential strong lens
+systems.  These will form the core of our training and testing datasets; our
+metric will be how well a CNN correctly identifies known and simulated lenses
+and non-lens systems.
 
@@
 From approximately 12000 fields of $440\times440\times3$ fields, we have
-constructed ______ cutouts sized $96\times96\times3$. We
+constructed approximately 30000 cutouts sized $96\times96\times3$. These cutouts
+are selected based on where citizen scientists clicked, on the theory that both
+'correct' and 'incorrect' selections provide useful information about the
+characteristics of gravitational lenses. In general, we have access to two
+broad classes of images: 'training' and 'test' images. The 'training' images
+include fields that were verified in advance to not contain any lenses as well
+as simulated lensed galaxies, quasars, and clusters. Many of the simulated
+objects are over-exaggerated and extremely obvious, but we also have access to
+a second 'refinement stage' of the project, where much harder simulations were
+given to users. The 'test' images are the fields that citizen scientists
+viewed, assessing whether a lens was in the field or not. In these 'test'
+images are 120 known strong gravitational lens systems, which are also included
+in this set. (The project confirms roughly half of these known lenses for
+reasonable definitions of completeness and purity.) For all the images we also
+have an associated probability that the project would evaluate that system as
+containing a lens.
 

### 2015-02-17 doc/Milestone/milestone.tex
milestone writing woohoo

--- 2015-02-17 doc/Milestone/milestone.tex
@@
 From approximately 12000 fields of $440\times440\times3$ fields, we have
-constructed ______ cutouts sized $96\times96\times3$. We
+constructed <> cutouts sized $96\times96\times3$. These cutouts
+
+We plan to train a classifier on this data using two different methods. First, we will code 
+our own convolutional net in python using theano. Second, we will apply transfer learning 
+techniques to train on a convolutional net galaxy morphology classifier, which has graciously 
+been made available to us by Ryan Keisler and which achieved 7th place in the 2014 Galaxy 
+Zoo Kaggle competition. This classifier runs $96\times96$ images through 3 convolutional 
+layers and 2 fully-connected layers and predicts a galaxy to have one of 37 enumerated 
+morphologies. We plan to train our classifier on top of the first full-connected layer, which 
+has 500 neurons.
 
@@
 \section{Intermediate/Preliminary Results}
+
+We have begun constructing our own convolutional network with two convolutional/max-pool 
+layers and two fully-connected layers. Although we have not yet trained on our lensing images,
+we can load these images and run them through our initialized classifier.
 

### 2015-02-17 doc/Milestone/milestone.tex

--- 2015-02-17 doc/Milestone/milestone.tex
@@
 
+We plan to train a classifier on this data using two different methods. First, we will code 
+our own convolutional net in python using theano. Second, we will apply transfer learning 
+techniques to train on a convolutional net galaxy morphology classifier, which has graciously 
+been made available to us by Ryan Keisler and which achieved 7th place in the 2014 Galaxy 
+Zoo Kaggle competition. This classifier runs $96\times96$ images through 3 convolutional 
+layers and 2 fully-connected layers and predicts a galaxy to have one of 37 enumerated 
+morphologies. We plan to train our classifier on top of the first full-connected layer, which 
+has 500 neurons.
+
 %-------------------------------------------------------------------------
 \section{Intermediate/Preliminary Results}
+
+We have begun constructing our own convolutional network with two convolutional/max-pool 
+layers and two fully-connected layers. Although we have not yet trained on our lensing images,
+we can load these images and run them through our initialized classifier.
 

### 2015-02-17 doc/Milestone/milestone.tex
milestone finished! close issue #3. Also a while back I could close issue #4

@@
 %\ifcvprfinal\pagestyle{empty}\fi
-\setcounter{page}{4321}
+\setcounter{page}{1}
 \begin{document}
@@
 poor completeness or poor purity -- and often poor both. Additionally, some
-algorithms are better at finding some types of lenses than others -- some
+algorithms are better at finding some types of lenses than others; some
 perform well on the brilliant arcs, but poorly on the multiply-imaged objects,
@@
 
-{\sc Space\,Warps}\xspace is a citizen-science initiative designed to overcome
-these two problems. The program has users examine images from the
-Canada-France-Hawaii Telescope Legacy Survey (CFHTLS) and vote on where they
-see lenses. Users are also assessed and trained with simulated lenses and known
-empty fields. By having thousands of users analyze a survey for short amounts
-of time each, it is hoped that a more complete sample of lenses can be
-discovered, which can then be fed into lens-finding algorithms to further
-improve their performance.
+{\sc Space\,Warps}\xspace (Marshall et. al, in prep.) is a citizen-science
+initiative designed to overcome these two problems. The program has users
+examine images from the Canada-France-Hawaii Telescope Legacy Survey (CFHTLS)
+and vote on where they see lenses. Users are also assessed and trained with
+simulated lenses and known empty fields. By having thousands of users analyze a
+survey for short amounts of time each, it is hoped that a more complete sample
+of lenses can be discovered, which can then be fed into lens-finding algorithms
+to further improve their performance.
+
+\begin{figure*}[!ht]
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/0.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/1.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/2.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/3.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+  \caption{Typical Space Warps duds. Insets indicate regions where volunteers typically
+  clicked.}
+  \label{fig:duds}
+\end{figure*}
+
+\begin{figure*}[!ht]
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/4.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/5.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/6.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=figs/gallery/7.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+  \caption{Typical sims. Insets indicate the location of the lens in the
+  image.}
+  \label{fig:sims}
+\end{figure*}
 
@@
 
-We plan to train a classifier on this data using two different methods. First, we will code 
-our own convolutional net in python using theano. Second, we will apply transfer learning 
-techniques to train on a convolutional net galaxy morphology classifier, which has graciously 
-been made available to us by Ryan Keisler and which achieved 7th place in the 2014 Galaxy 
-Zoo Kaggle competition. This classifier runs $96\times96$ images through 3 convolutional 
-layers and 2 fully-connected layers and predicts a galaxy to have one of 37 enumerated 
-morphologies. We plan to train our classifier on top of the first full-connected layer, which 
-has 500 neurons.
+It is clear that we do not have enough data. Luckily, we also know that our
+lens objects must obey certain symmetry properties, so it is quite easy to
+augment our data. For example, we know that strong lens systems should be
+independent of rotations as well as small amounts of stretching and
+translation, so our data can be augmented by applying those transformations to
+our images.
+
+We plan to train a classifier on this data using two different methods. First,
+we will code our own convolutional net in python using theano. Second, we will
+apply transfer learning techniques to train on a convolutional net galaxy
+morphology classifier, which has graciously been made available to us by Ryan
+Keisler and which achieved 7th place in the 2014 Galaxy Zoo Kaggle competition.
+This classifier runs $96\times96$ images through 3 convolutional layers and 2
+fully-connected layers and predicts a galaxy to have one of 37 enumerated
+morphologies. We plan to train our classifier on top of the first
+fully-connected layer, which has 500 neurons.
 
@@
 
-We have begun constructing our own convolutional network with two convolutional/max-pool 
-layers and two fully-connected layers. Although we have not yet trained on our lensing images,
-we can load these images and run them through our initialized classifier.
+We have created the cutouts of our images from the catalogs.
+We have begun constructing our own convolutional network with two
+convolutional/max-pool layers and two fully-connected layers. Although we have
+not yet trained on our images, we can load these images and run them
+through our initialized classifier. Figures~\ref{fig:duds} and
+\ref{fig:sims}\xspace show example fields with cutouts inlaid.
 

### 2015-03-08 code/analyze_TL_database.py
push for 8 march

@@ -6,6 +6,8 @@ Methods to apply:
 
 TODO: apply some kind of cut
 TODO: mean subtraction and std division
+TODO: save!
+TODO: train test split on cutouts, not the data augmented
 """
 
 import pandas as pd
@@ -17,12 +19,20 @@ from sklearn.metrics import roc_curve
 import matplotlib.pyplot as plt
 
 # this is something like 4.5 gigs of memory!
-databasepath = '/Users/cpd/Desktop/batches/database.csv'
-data_all = pd.read_csv(databasepath)
+project_dir = '/Users/cpd/Projects/strongcnn/'
+datapath = '/Users/cpd/Desktop/batches/'
+
+data = pd.read_csv(datapath + 'database.csv',
+                       index_col=522)
+
+# load up catalog for cutouts
+cat = pd.read_csv(project_dir + 'catalog/cluster_catalog.csv')
+
 print('data loaded')
 # we also need to filter out entries who are of type test which are
 # indeterminate
-data = data_all[data_all['category'] == 'training']
+data = data[data['category'] == 'training']
+cat = cat[cat['category'] == 'training']
 
 # extract features
 feature_cols = ['nn{0}'.format(i) for i in xrange(500)]
@@ -33,14 +43,40 @@ X = data[feature_cols].values
 y = data['alpha'] == 1
 
 print(sum(y))
-
 print(X.shape)
+
+
+# get cuts on cutoutname for stage1, stage2, and all
+cut_train, cut_test = train_test_split(cat['cutoutname'].values, test_size=0.2)
+cut_train_1, cut_test_1 = train_test_split(cat[cat['stage'] == 1]['cutoutname'].values, test_size=0.2)
+cut_train_2, cut_test_2 = train_test_split(cat[cat['stage'] == 2]['cutoutname'].values, test_size=0.2)
+
+"""
+Need:
+    X_test, X_train
+    y_test, y_train
+for all, _1, _2
+"""
+X_train = data.loc[cut_train][feature_cols].values
+y_train = data.loc[cut_train]['alpha'] == 1
+X_test = data.loc[cut_test][feature_cols].values
+y_test = data.loc[cut_test]['alpha'] == 1
[…]

### 2015-03-11 doc/Poster/example.tex
modified analysis. Made poster

% An example document showing the use of the ucbposter, based on the
% infposter demo developed by M.Reddy (MXR), 24/2/97 and updated by
% Mary E Foster, 10/5/2002.

\documentclass{ucbposter}

\let\AuthorColour\red

%http://www.inf.ed.ac.uk/systems/tex/infposter/infposter.html
%\epsfigure[<width-percent>]{<filename>}{<caption>}

\begin{document}

\title{An Example of the UCB Poster Class}
%\subtitle{A simple title for a simple poster.}
\author{Your Name(s)}
\email{an@address}
%\homepage{http://radlab.cs.berkeley.edu/}
\makeposter

\section{Poster content}

\begin{itemize}
  \item An example bullet point
  \begin{itemize}
    \item Example
  \end{itemize}
\end{itemize}

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.
\end{document}

### 2015-03-11 doc/Poster/poster.tex
modified analysis. Made poster

% An example document showing the use of the ucbposter, based on the
% infposter demo developed by M.Reddy (MXR), 24/2/97 and updated by
% Mary E Foster, 10/5/2002.

\documentclass{ucbposter}

\let\AuthorColour\red

%http://www.inf.ed.ac.uk/systems/tex/infposter/infposter.html
%\epsfigure[<width-percent>]{<filename>}{<caption>}
% \begin{itemize}
%   \item An example bullet point
%   \begin{itemize}
%     \item Example
%   \end{itemize}
% \end{itemize}

\begin{document}

\title{Recognizing Strong Gravitational Lenses}
\author{Chris Davis and Andrew McLeod}
\email{cpd@stanford.edu}
%\homepage{http://radlab.cs.berkeley.edu/}
\makeposter

\section{Strong Gravitational Lenses}

\epsfigure[40]{Figures/lens.png}{Example
strong gravitational lens discovered by \textsc{Space Warps}.}

\begin{itemize}
\item{Strong lenses -- systems with possibly multiple magnified and
distorted images of background objects (sources) due to the deflection of light
by massive foreground objects (lenses) -- can be used as astrophysical tools to
probe mass distributions, magnify distant objects, and measure fundamental
cosmological parameters.
\item{The main problem with strong gravitational lenses is their rarity:
  \begin{itemize}
    \item{For
modern wide-field optical surveys which observe tens of millions of galaxies
over many thousand square degrees, one expects to find only several hundred
such systems.} \item{While strong lenses
are relatively easy to spot by eye, machine learning techniques have thus far
been unsuccessful in reliably identifying strong lenses and distinguishing them from
image artifacts (e.g. cosmic rays) and other false positives (e.g. pixel
noise). Maybe CNNs will help!}
\end{itemize}}
\end{itemize}
\epsfigure[90]{Figures/sims.png}{Example fields (with cutouts in corner)
containing simulated lenses.}

\epsfigure[90]{Figures/duds.png}{Example fields (with cutouts in corner)
containing no lenses. Cutouts indicate regions used as inputs for `duds'.}

\section{Space Warps Dataset}

\begin{itemize}
  \item{\textsc{Space Warps} (\url{www.spacewarps.com}) is an online crowd-sourced gravitational
    lens detection system that invites citizen
scientists to interpret real data from the Canada-France-Hawaii Telescope
Legacy Survey.}
\item{Volunteers also analyze artificial ``training'' images of simulated
    lenses as well as images confirmed to have no strong gravitational lenses.
    Simulated lenses are also further demarcated by lens type (lensed quasar,
  lensing galaxy, lensing cluster), each of which has different potential
cosmological applications.}
\item{\textsc{Space Warps} is divided into two stages, where the second stage
    refines the results of the first by using considerably more difficult
  training images. There are 24177 training images in Stage 1 and 1876 in Stage
2. We also have 9030 ``test'' images (cutouts whose status is not known but for
which we are interested in making reasonable guesses).}
\item{\textsc{Space Warps} will provide a ``curated'' dataset of gravitational
  lenses and lens-look-alikes to facilitate the training of future automated detection
algorithms.}
\item{We train our networks on the ``training'' images only, of which there are
  20743 ``duds'' and 5310 ``sims''. From fields of 440 x 440 x 3, we cut
  out 96 x 96 x 3 stamps based on the locations \textsc{Space Warps}
users clicked.}
\end{itemize}

\section{Our Own CNN}

\section{Transfer Learning with Galaxy Morphology}

%\epsfigure[40]{Figures/architecture.pdf}{Architecture of the convolution
%  neural network trained on the galaxy zoo morphologies. The feature vector we
%use comes from the fully connected layer.}
%
%\epsfigure[40]{Figures/roc_curve.pdf}{Receiver Operating Curves for the Random
%  Forest Transfer Learning and the \textsc{Space Warps} systems, divided by
%  stage.}

  \epsfigure[90]{Figures/arch_roc.png}{Architecture of the convolution
  neural network trained on the galaxy zoo morphologies and Receiver Operating Curves for the Random
  Forest Transfer Learning and the \textsc{Space Warps} systems, divided by
  stage. }

\begin{itemize}
  \item{Kaggle held a competition to classify galaxy morphology using galaxies
    from the Sloan Digital Sky Survey, a comparable but older survey to the
  Canada-France-Hawaii-Telescope. Many of the high-ranking competitors used
convolution neural networks.}
  \item{We examined how well we could transfer the learning of morphology to
    identifying strong gravitational lenses by taking a convolution neural
  network trained on the Kaggle Galaxy Zoo data by Ryan Keisler and running our cutouts through
to obtain a length 500 feature vector.}
  \item{We then trained these feature vectors on a random forest, which is an
    ensemble of decision trees wherein the order and number of feature vectors
  used in a given tree is random.}
\item{For comparison, we plot the results from citizen scientists on these two
  stages as well. While we do worse overall, our results were extremely quick
 to compute, while the \textsc{Space Warps} project involved the
efforts of tens of thousands over several months.}
\item{Nevertheless, our training results take citizen-scientist inputs. There
  is much potential in the future for machines and citizen-scientists to learn
from one another!}
\end{itemize}

\section{Future}

\begin{itemize}
  \item{Expand categorization procedure beyond binary lens-not to the specific
    types of lenses.}
  \item{Explore deeper types of transfer learning (e.g. retraining parts of the
    network).}
  \item{SOMETHING ELSE}
\end{itemize}

\section{Acknowledgements}

We received access to the dataset through Phil Marshall
(\url{pjm@slac.stanford.edu}), who also graciously explained to us how
\textsc{Space Warps}
currently works. The convolutional neural network upon which the transfer
learning is based was kindly provided to us by Ryan Keisler
(\url{rkeisler@stanford.edu}).
All errors in interpretation or otherwise are entirely our own.

\end{document}

### 2015-03-11 code/analyze_TL_database.py

@@ -17,6 +17,7 @@ from sklearn.ensemble import RandomForestClassifier
 from sklearn.cross_validation import cross_val_score, train_test_split
 from sklearn.metrics import roc_curve
 import matplotlib.pyplot as plt
+from sklearn.preprocessing import label_binarize
 
 # this is something like 4.5 gigs of memory!
 project_dir = '/Users/cpd/Projects/strongcnn/'
@@ -70,11 +71,14 @@ y_train_2 = data[data['stage'] == 2].loc[cut_train_2]['alpha'] == 1
 X_test_2 = data[data['stage'] == 2].loc[cut_test_2][feature_cols].values
 y_test_2 = data[data['stage'] == 2].loc[cut_test_2]['alpha'] == 1
 
-import ipdb; ipdb.set_trace()
+mult_dict = {'dud': 0, 'lensed quasar': 1, 'lensing cluster': 2, 'lensed galaxy': 3}
+# y_train_mult = np.array([y_train[i] * mult_dict[data.loc[cut_train]['flavor'][i]] for i in xrange(len(y_train))])
+
+
 # train on svm and rf
-clf_rf = RandomForestClassifier()
-#scores_rf = cross_val_score(clf_rf, X, y)
-#print('rf: ', scores_rf)
+clf_rf = RandomForestClassifier(n_estimators=50, n_jobs=-1)
+scores_rf = cross_val_score(clf_rf, X, y)
+print('rf: ', scores_rf)
 
 
 clf_rf.fit(X_train, y_train)
@@ -90,18 +94,18 @@ plt.figure()
 plt.xlabel('Feature Number')
 plt.ylabel('Feature Importance')
 plt.plot(clf_rf.feature_importances_, 'k--')
-
+plt.savefig(project_dir + 'doc/Poster/Figures/feature_importance.pdf')
 
 # repeat for stage1 and stage 2
-clf_rf_1 = RandomForestClassifier()
+clf_rf_1 = RandomForestClassifier(n_estimators=50, n_jobs=-1)
 clf_rf_1.fit(X_train_1, y_train_1)
-y_score_1 =  clf_rf.predict_proba(X_test_1)
+y_score_1 =  clf_rf_1.predict_proba(X_test_1)
 y_score_1 = y_score_1[:, 1]
[…]

### 2015-03-11 doc/Poster/poster.tex
latest (and final) version of the poster

@@
 \author{Chris Davis and Andrew McLeod}
-\email{cpd@stanford.edu}
+\email{cpd@stanford.edu 
+ajmcleod@stanford.edu}
 %\homepage{http://radlab.cs.berkeley.edu/}
@@
 \section{Strong Gravitational Lenses}
-
 \epsfigure[40]{Figures/lens.png}{Example
 strong gravitational lens discovered by \textsc{Space Warps}.}
-
 \begin{itemize}
-\item{Strong lenses -- systems with possibly multiple magnified and
+\item{Strong lenses -- systems with magnified and
 distorted images of background objects (sources) due to the deflection of light
@@
 probe mass distributions, magnify distant objects, and measure fundamental
-cosmological parameters.
-\item{The main problem with strong gravitational lenses is their rarity:
+cosmological parameters.}
+\item{The main problem with strong gravitational lenses is their rarity:}
   \begin{itemize}
@@
 over many thousand square degrees, one expects to find only several hundred
-such systems.} \item{While strong lenses
+such systems.} 
+\item{While strong lenses
 are relatively easy to spot by eye, machine learning techniques have thus far
 been unsuccessful in reliably identifying strong lenses and distinguishing them from
-image artifacts (e.g. cosmic rays) and other false positives (e.g. pixel
-noise). Maybe CNNs will help!}
-\end{itemize}}
+image artifacts (e.g. cosmic rays) and other false positives. Maybe CNNs will help!}
 \end{itemize}
+\end{itemize}
+
+%%% second pane
+
 \epsfigure[90]{Figures/sims.png}{Example fields (with cutouts in corner)
@@
 \section{Space Warps Dataset}
-
 \begin{itemize}
@@
     lenses as well as images confirmed to have no strong gravitational lenses.
-    Simulated lenses are also further demarcated by lens type (lensed quasar,
+    Simulated lenses are further demarcated by lens type (lensed quasar,
   lensing galaxy, lensing cluster), each of which has different potential
@@
 which we are interested in making reasonable guesses).}
-\item{\textsc{Space Warps} will provide a ``curated'' dataset of gravitational
-  lenses and lens-look-alikes to facilitate the training of future automated detection
+\item{\textsc{Space Warps} will provide a curated dataset of gravitational
+  lenses and lens look-alikes to facilitate the training of future automated detection
 algorithms.}
@@
   20743 ``duds'' and 5310 ``sims''. From fields of 440 x 440 x 3, we cut
-  out 96 x 96 x 3 stamps based on the locations \textsc{Space Warps}
-users clicked.}
+  out 96 x 96 x 3 stamps based on \textsc{Space Warps} inputs.}
 \end{itemize}
 
-\section{Our Own CNN}
+%%% third pane
 
-\section{Transfer Learning with Galaxy Morphology}
+\section{Transfer Learning with Galaxy Zoo}
 
@@
 
-  \epsfigure[90]{Figures/arch_roc.png}{Architecture of the convolution
-  neural network trained on the galaxy zoo morphologies and Receiver Operating Curves for the Random
-  Forest Transfer Learning and the \textsc{Space Warps} systems, divided by
-  stage. }
-
 \begin{itemize}
   \item{Kaggle held a competition to classify galaxy morphology using galaxies
-    from the Sloan Digital Sky Survey, a comparable but older survey to the
-  Canada-France-Hawaii-Telescope. Many of the high-ranking competitors used
-convolution neural networks.}
+    from the Sloan Digital Sky Survey, a comparable (but older) survey to the
+  Canada-France-Hawaii-Telescope.}
   \item{We examined how well we could transfer the learning of morphology to
@@
   \item{We then trained these feature vectors on a random forest, which is an
-    ensemble of decision trees wherein the order and number of feature vectors
+    ensemble of decision trees wherein the order and feature vectors
   used in a given tree is random.}
@@
 efforts of tens of thousands over several months.}
-\item{Nevertheless, our training results take citizen-scientist inputs. There
-  is much potential in the future for machines and citizen-scientists to learn
-from one another!}
+\end{itemize}
+
+  \epsfigure[90]{Figures/arch_roc.png}{
+    Architecture of the convolution neural network trained on the galaxy zoo
+    morphologies and Receiver Operating Curves for the Random Forest Transfer
+    Learning and the \textsc{Space Warps} systems. }
+
+\section{Convolution Neural Network}
+\begin{itemize}
+  \item{Using \textsc{theano}, we have built a 5-layer CNN which we are
+    training from scratch on the \textsc{Space Warps} Dataset.}
+\item{Our algorithm uses RMSprop and dropout at each layer, and has an
+  architecture of:}
+\begin{itemize}
+\item{Convolution Layer of stride 4 and depth 10}
+\item{Max Pool Layer of stride 2}
+\item{Convolution Layer of stride 4 and depth 10}
+\item{Max Pool Layer of stride 2}
+\item{Fully Connected Layer with 50 neurons}
+\end{itemize}
+\item{Our classifier has been tested and works, but more training epochs are
+  needed before it will be competitive with our transfer learning results.}
 \end{itemize}
@@
 \section{Future}
-
 \begin{itemize}
-  \item{Expand categorization procedure beyond binary lens-not to the specific
-    types of lenses.}
-  \item{Explore deeper types of transfer learning (e.g. retraining parts of the
+  \item{Explore deeper transfer learning (e.g. retraining parts of the
     network).}
-  \item{SOMETHING ELSE}
+  \item{Continue training our own CNN on the Space Warps Dataset directly.}
 \end{itemize}
 
-\section{Acknowledgements}
-
+\subsection{Acknowledgements}
 We received access to the dataset through Phil Marshall

### 2015-03-14 doc/Course_Paper/paper.tex
adding course paper doc

--- 2015-02-17 doc/Milestone/milestone.tex
@@
 
-% %%%%%%%%% ABSTRACT
-% \begin{abstract}
-%    The ABSTRACT is to be in fully-justified italicized text, at the top
-%    of the left-hand column, below the author and affiliation
-%    information. Use the word ``Abstract'' as the title, in 12-point
-%    Times, boldface type, centered relative to the column, initially
-%    capitalized. The abstract is to be in 10-point, single-spaced type.
-%    Leave two blank lines after the Abstract, then begin the main text.
-%    Look at previous CVPR abstracts to get a feel for style and length.
-% \end{abstract}
+%%%%%%%%% ABSTRACT
+\begin{abstract}
+    The detection of strong gravitational lenses
+\end{abstract}
 
@@
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/0.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/0.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/1.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/1.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/2.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/2.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/3.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/3.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
@@
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/4.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/4.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/5.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/5.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/6.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/6.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=figs/gallery/7.png,width=\linewidth,angle=0,clip=}
+\centering\epsfig{file=Figures/gallery/7.png,width=\linewidth,angle=0,clip=}
 \end{minipage} \hfill
@@
 We plan to train a classifier on this data using two different methods. First,
-we will code our own convolutional net in python using theano. Second, we will
-apply transfer learning techniques to train on a convolutional net galaxy
-morphology classifier, which has graciously been made available to us by Ryan
-Keisler and which achieved 7th place in the 2014 Galaxy Zoo Kaggle competition.
-This classifier runs $96\times96$ images through 3 convolutional layers and 2
-fully-connected layers and predicts a galaxy to have one of 37 enumerated
-morphologies. We plan to train our classifier on top of the first
+we will code our own convolutional net in python using \textsc{theano}. Second,
+we will apply transfer learning techniques to train on a convolutional net
+galaxy morphology classifier, which has graciously been made available to us by
+Ryan Keisler and which achieved 7th place in the 2014 Galaxy Zoo Kaggle
+competition.  This classifier runs $96\times96$ images through 3 convolutional
+layers and 2 fully-connected layers and predicts a galaxy to have one of 37
+enumerated morphologies. We plan to train our classifier on top of the first
 fully-connected layer, which has 500 neurons.
+
+%-------------------------------------------------------------------------
+\section{Results}
+
+% Not sure if we put anything here?
+
+%-------------------------------------------------------------------------
+\subsection{Convolution Neural Network}
+
+% TODO: Insert figure for architecture
+
+%-------------------------------------------------------------------------
+\subsection{Transfer Learning}
+
+% TODO: Touch up description
+\begin{figure}%[!ht]
+\centering
+\epsfig{file=Figures/architecture.pdf,width=\linewidth,angle=0,clip=}
+\caption{Architecture of the convolution
+  neural network trained on the galaxy zoo morphologies. The feature vector we
+use comes from the fully connected layer.}
+\label{fig:architecture}
+\end{figure}
+
+% TODO: Redescribe the network, including data augmentation and such.
+
+%-------------------------------------------------------------------------
+\section{Discussion}
+
+% TODO: Remake with linear classifiers.
+% TODO: Include results from CNN as well.
+% TODO: Touch up description
+\begin{figure}%[!ht]
+\centering
+\epsfig{file=Figures/roc_curve.pdf,width=\linewidth,angle=0,clip=}
+\caption{Receiver Operating Curves for the Random
+  Forest Transfer Learning and the \textsc{Space Warps} systems, divided by
+  stage.}
+  \label{fig:roc}
+\end{figure}
+
+%-------------------------------------------------------------------------
+\section{Conclusions}
+
+% Discuss how well things went. Or didn't go.
+
+%-------------------------------------------------------------------------
+\section{Acknowledgements}
+We received access to the dataset through Phil Marshall
+(\url{pjm@slac.stanford.edu}), who also graciously explained to us how
+\textsc{Space Warps}
+currently works. The convolutional neural network upon which the transfer
+learning is based was kindly provided to us by Ryan Keisler
+(\url{rkeisler@stanford.edu}).
+All errors in interpretation or otherwise are entirely our own.
 
@@
 
-% {\small
-% \bibliographystyle{ieee}
-% \bibliography{egbib}
-% }
+% TODO: copy relevant bib stuff over so I don't have a 4 mb bib
+{\small
+\bibliographystyle{ieee}
+\bibliography{coursebib}
+}
 

### 2015-03-15 doc/Course_Paper/paper.tex
updating course paper

@@
 \begin{abstract}
-    The detection of strong gravitational lenses
+    The detection of a large, representative set of strong gravitational lenses
+    could greatly aid in our understanding of cosmology. Unfortunately they are
+    quite rare, and the best techniques now revolve around squads of scientists
+    manually scanning through images. This is presently borderline
+    unsustainable and will be laughably inefficient with the advent of the
+    Large Synoptic Survey Telescope. Here we examine the effectiveness of
+    convolution neural networks and transfer learning for automated detection
+    algorithms of strong gravitational lenses. We use images from the
+    \textsc{Space Warps} project, a citizen science initiative to examine tens
+    of thousands of fields of galaxies for the presence of strong gravitational
+    lenses. 
+    % TODO: Insert results!
+    Scripts used in the analysis of this paper are freely available at
+    \url{https://github.com/cpadavis/strongcnn}. The images are currently only
+    available to those who contact the author, but will be available to the
+    public in the near future.
 \end{abstract}
@@
 \section{Introduction}
+
+% TODO: insert references!
 
@@
 passing through a deep gravitational potential are termed strong gravitational
-lenses.
+lenses. The gravitational lensing signal can heuristically be thought of as a
+trade off between a couple factors: larger gradients in the gravitational
+potential create larger distortions (and larger gradients in the gravitational
+potential tend to reside near to the center of the foreground galaxy or galaxy
+cluster); the more separated the foreground and background objects are, the
+bigger the proportional distortion on the background object by the foreground
+(but the farther away the background object is, the smaller it
+appears\footnote{Note that this is only true in the ``low redshift'' universe:
+  when objects are farther away than a cosmological redshift of $z \approx 2.5$
+or about 2.5 Gyr after the birth of the Universe, they will actually
+\textit{grow} in angular extent.}).
 
@@
 \begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=Figures/gallery/0.png,width=\linewidth,angle=0,clip=}
-\end{minipage} \hfill
-\begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=Figures/gallery/1.png,width=\linewidth,angle=0,clip=}
-\end{minipage} \hfill
-\begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=Figures/gallery/2.png,width=\linewidth,angle=0,clip=}
-\end{minipage} \hfill
-\begin{minipage}[b]{0.24\linewidth}
-\centering\epsfig{file=Figures/gallery/3.png,width=\linewidth,angle=0,clip=}
-\end{minipage} \hfill
-  \caption{Typical Space Warps duds. Insets indicate regions where volunteers typically
-  clicked.}
-  \label{fig:duds}
-\end{figure*}
-
-\begin{figure*}[!ht]
-\begin{minipage}[b]{0.24\linewidth}
 \centering\epsfig{file=Figures/gallery/4.png,width=\linewidth,angle=0,clip=}
@@
   \caption{Typical sims. Insets indicate the location of the lens in the
-  image.}
+  image. These insets are fed into our training system.}
   \label{fig:sims}
+\end{figure*}
+
+\begin{figure*}[!ht]
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=Figures/gallery/0.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=Figures/gallery/1.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=Figures/gallery/2.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.24\linewidth}
+\centering\epsfig{file=Figures/gallery/3.png,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+  \caption{Typical Space Warps duds. Insets indicate regions where volunteers
+  typically clicked. These insets are then fed into our training system.}
+  \label{fig:duds}
 \end{figure*}
@@
      images?}
-\item{ What processing needs to be done on the data? What kind of scaling of
-       pixel data is appropriate for automated detection? Should we compress
-     the five different 'colors' to a smaller number of dimensions?}
 \item{ How do citizen-scientists do compared with this automated system?}
-   \item{Can we use the results of citizen-scientists to train the CNN?}
-\item{What sets of classifications are needed? Are we better served sticking
-       to 'lens' and 'not lens', or should we use several classification
-       categories ('lensed arcs', 'lensed multiple images', 'non-lens pixel
-     noise')?}
+\item{ Can we use the results of citizen-scientists to train the CNN?}
+\item{ How well does using features extracted from a convolution neural network
+       trained on galaxy morphology perform when determining the presence of
+     strong gravitational lenses?}
 \end{itemize}
@@
 
-We plan to train a classifier on this data using two different methods. First,
-we will code our own convolutional net in python using \textsc{theano}. Second,
-we will apply transfer learning techniques to train on a convolutional net
+We train a classifier on this data using two different methods. First,
+we code our own convolutional net in python using \textsc{theano}. Second,
+we apply transfer learning techniques to train on a convolutional net
 galaxy morphology classifier, which has graciously been made available to us by
@@
 layers and 2 fully-connected layers and predicts a galaxy to have one of 37
-enumerated morphologies. We plan to train our classifier on top of the first
-fully-connected layer, which has 500 neurons.
+enumerated morphologies. (See Figure~\ref{fig:ryan_architecture}.) We
+train classifiers on top of the first fully-connected layer, which has 500
+neurons.
+
+%-------------------------------------------------------------------------
+\subsection{The \textsc{Space Warps} Catalog}
+
+\textsc{Space Warps} is a web-based service that enables the discovery of
+strong gravitational lenses in wide-field imaging surveys by large numbers of
+people. Carefully produced color composite images are displayed to volunteers
+via a classification interface which records their estimates of the positions
+of candidate lensed features. Simulated lenses, and expert-classified
+non-lenses, are inserted into the image stream at random intervals; this
+training set is used to give the volunteers feedback on their performance, and
+to estimate a dynamically-updated probability for any given image to contain a
+lens. Low probability systems are retired from the site periodically,
+concentrating the sample towards a set of candidates; this ``stage 1'' set is
+then re-classified by the volunteers in a second refinement stage. This ``stage
+2'' has a different set of training images, ones that are generally considered
+`harder'. Most stage 1 simulated lenses are very obvious\footnote{Very bright
+and blue quasars multiply-imaged around a small red galaxy, very bright,
+separated, and full \href{http://en.wikipedia.org/wiki/Einstein_ring}{Einstein
+rings.}}, while simulated lenses in stage 2 are often much more subtle.
+\footnote{Dim multiply-imaged quasars of varying magnitude, dim and incomplete
+Einstein rings located close to a galaxy.} Figures~\ref{fig:duds} and
+\ref{fig:sims}\xspace show example stage 2 fields with cutouts inlaid. Notice
+that while the first three images in Figure~\ref{fig:sims} are very clearly
+strong gravitational lenses \footnote{For the neophyte: the first is a broken
+  blue arc around a central red galaxy; the second is a blue arc around a
+  central yellow galaxy; the third is a multiply-imaged blue quasar (images
+appear above and below the central galaxy), the fourth is a broken dim arc
+located behind a very bright foreground galaxy.}, the fourth is very difficult
+to find.  Unfortunately, we would very much like to find these, because there
+are many such systems and they contain important information about the mass
+structures at the centers of galaxies.\footnote{The trade off is this: the rate
+  of alignment between foreground and background objects decreases as one
+  decreases the area around a foreground object, but the strength of strong
+  gravitational distortions -- and the signal we can pull out from identifying
+  such systems -- increases as one gets closer to the center of the foreground
+  object. A yet further complication to this is that background objects are
+  naturally fainter than foreground objects, \textit{and} the foreground
+  objects with the highest gravitational potential (and hence the largest
+distortions of background images) tend to also be the brightest objects. Both
+these complications render the task even more difficult.}
+Figure~\ref{fig:duds} highlights the difficulties of this task. Each `dud' has
+features that conceivably look like strong gravitational lensing, but are in
+actuality some other confounding effect: color gradients from variations in the
+Point Spread Function between the different color bands, dust surrounding a
+galaxy, galaxies that are actually in the same cluster, and chance alignments
+of background galaxies and foreground stars.
+
+The fields users observe are $440\times440$ size images, containing multiple
+potential locations for strong gravitational lenses, although it is unlikely
+that a field contains more than one strong gravitational lens. In order to
+generate $96\times96$ images of lenses and non-lenses, we use the recorded
+estimates of the positions of candidate lensed feature. More specifically, we
+apply the \textsc{DBSCAN} clustering algorithm, which agglomeratively grows
+clusters such that that are within a minimum distance and contain a minimum
+number of samples. \textsc{DBSCAN} is a convenient choice of a clustering
+algorithm because it has a well-defined way of rejecting outliers, which we
+generically interpret as genuine ``mis-clicks'' on the part of users. We are
+very generous in the definition of a cluster and only require two members within
+100 pixels of each other to form a cluster.
+
+The fear of noisy clusters is this: the real task of our techniques is to
+distinguish strong gravitational lens systems from other configurations of
+galaxies (for example, random alignments of galaxies). Noisy clicks end up
+creating random cutouts of the field, slightly changing the task of our
+classifier to distinguishing strong gravitational lens systems from random
+cutouts from the field of a galaxy survey.  Inclusion of noise however ended up
+not being an issue for stage 2: non-lenses, where the correct action on the
+part of the user is to leave no marker, have a median of 28 markers in stage 2,
+while simulated lenses (where the correct answer is to click at a specific
+location) have a median of 180 markers. In stage 1, the simulated lenses have a
+median of 80 markers, while the duds have a median of 3 markers (and a mean of
+9.6). It may be the case that noise is injected in the stage 1 non-lens sample.
+
+Overall, our base dataset has 24,177 images from stage 1, of which 5159 are of
+simulated lenses, and 1876 images from stage 2, of which 151 are simulated
+lenses. We also have 9030 classifications that stage 2 users made of images in
+the CFHTLS survey where it is unknown whether they contain a lens or not. From
+these classifications, a list of approximately 40 candidate strong
+gravitational lensing objects have been found which will soon receive
+spectroscopic follow-up to confirm whether they are strong gravitational
+lensing systems or not.\footnote{Spectroscopy can yield precise redshifts of
+  different objects in a field. This way, if different parts of a strong
+  gravitational lens arc are at the same redshift, or if the multiply-imaged
+quasars are, then one can confirm that we are really seeing such a system.} A
+future project with this work would be to link the probabilities from the
+\textsc{Space Warps} system with the probabilities obtained by a detection
+algorithm.
 
@@
 use comes from the fully connected layer.}
-\label{fig:architecture}
+\label{fig:ryan_architecture}
 \end{figure}
 
+\begin{figure*}%[!ht] 
+\begin{minipage}[b]{0.48\linewidth}
+\centering\epsfig{file=Figures/roc_curve_stage_1.pdf,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+\begin{minipage}[b]{0.48\linewidth}
+\centering\epsfig{file=Figures/roc_curve_stage_2.pdf,width=\linewidth,angle=0,clip=}
+\end{minipage} \hfill
+  \caption{Receiver Operating Curves the \textsc{Space Warps} system and
+    different linear classifiers trained on feature vectors extracted from a
+    convolution neural network originally used to determine galaxy
+  morphologies. We find that of the linear classifiers on the feature vectors,
+support vector machine classifications perform best on the test dataset,
+however all the feature vectors perform worse than the users themselves. Note
+that the $x$-axis stops at a false positive rate of 0.5, and the $y$-axis
+begins at a true positive rate of 0.5.}
+\label{fig:roc_tl} 
+\end{figure*}
+
 % TODO: Redescribe the network, including data augmentation and such.
+Transfer Learning provides a mechanism by which
 
@@
 
+% TODO: make figure with probability distributions between classifiers and sw
+
 %-------------------------------------------------------------------------
@@
 We received access to the dataset through Phil Marshall
-(\url{pjm@slac.stanford.edu}), who also graciously explained to us how
+\footnote{\url{pjm@slac.stanford.edu}}, who also graciously explained to us how
 \textsc{Space Warps}
@@
 learning is based was kindly provided to us by Ryan Keisler
-(\url{rkeisler@stanford.edu}).
+\footnote{\url{rkeisler@stanford.edu}}.
 All errors in interpretation or otherwise are entirely our own.
@@
 not yet trained on our images, we can load these images and run them
-through our initialized classifier. Figures~\ref{fig:duds} and
-\ref{fig:sims}\xspace show example fields with cutouts inlaid.
+through our initialized classifier.
 

### 2015-03-15 code/analyze_TL_database.py

@@ -18,6 +18,9 @@ from sklearn.cross_validation import cross_val_score, train_test_split
 from sklearn.metrics import roc_curve
 import matplotlib.pyplot as plt
 from sklearn.preprocessing import label_binarize
+from sklearn.linear_model import SGDClassifier
+from scipy.stats import randint as sp_randint  # for searching for best hyperparameters!
+from sklearn.grid_search import GridSearchCV
 
 # this is something like 4.5 gigs of memory!
 project_dir = '/Users/cpd/Projects/strongcnn/'
@@ -58,56 +61,86 @@ Need:
     y_test, y_train
 for all, _1, _2
 """
-X_train = data.loc[cut_train][feature_cols].values
-y_train = data.loc[cut_train]['alpha'] == 1
-X_test = data.loc[cut_test][feature_cols].values
-y_test = data.loc[cut_test]['alpha'] == 1
-X_train_1 = data[data['stage'] == 1].loc[cut_train_1][feature_cols].values
-y_train_1 = data[data['stage'] == 1].loc[cut_train_1]['alpha'] == 1
-X_test_1 = data[data['stage'] == 1].loc[cut_test_1][feature_cols].values
-y_test_1 = data[data['stage'] == 1].loc[cut_test_1]['alpha'] == 1
-X_train_2 = data[data['stage'] == 2].loc[cut_train_2][feature_cols].values
-y_train_2 = data[data['stage'] == 2].loc[cut_train_2]['alpha'] == 1
-X_test_2 = data[data['stage'] == 2].loc[cut_test_2][feature_cols].values
-y_test_2 = data[data['stage'] == 2].loc[cut_test_2]['alpha'] == 1
+
+cuts = {'all': (cut_train, cut_test),
+        1: (cut_train_1, cut_test_1),
+        2: (cut_train_2, cut_test_2),}
+
+train = {# 'all': (data.loc[cut_train][feature_cols].values,
+         #         data.loc[cut_train]['alpha'] == 1),
+         1: (data[data['stage'] == 1].loc[cut_train_1][feature_cols].values,
+             data[data['stage'] == 1].loc[cut_train_1]['alpha'] == 1),
+         2: (data[data['stage'] == 2].loc[cut_train_2][feature_cols].values,
+             data[data['stage'] == 2].loc[cut_train_2]['alpha'] == 1),
+        }
+test = {# 'all': (data.loc[cut_test][feature_cols].values,
[…]

### 2015-03-15 code/cs231n/classifiers/convnet.py

@@ -1,0 +1,419 @@
+import numpy as np
+
+from cs231n.layers import *
+from cs231n.fast_layers import *
+from cs231n.layer_utils import *
+
+
+def two_layer_convnet(X, model, y=None, reg=0.0, dropout=1.0):
+  """
+  Compute the loss and gradient for a simple two-layer ConvNet. The architecture
+  is conv-relu-pool-affine-softmax, where the conv layer uses stride-1 "same"
+  convolutions to preserve the input size; the pool layer uses non-overlapping
+  2x2 pooling regions. We use L2 regularization on both the convolutional layer
+  weights and the affine layer weights.
+
+  Inputs:
+  - X: Input data, of shape (N, C, H, W)
+  - model: Dictionary mapping parameter names to parameters. A two-layer Convnet
+    expects the model to have the following parameters:
+    - W1, b1: Weights and biases for the convolutional layer
+    - W2, b2: Weights and biases for the affine layer
+  - y: Vector of labels of shape (N,). y[i] gives the label for the point X[i].
+  - reg: Regularization strength.
+
+  Returns:
+  If y is None, then returns:
+  - scores: Matrix of scores, where scores[i, c] is the classification score for
+    the ith input and class c.
+
+  If y is not None, then returns a tuple of:
+  - loss: Scalar value giving the loss.
+  - grads: Dictionary with the same keys as model, mapping parameter names to
+    their gradients.
+  """
+  
+  # Unpack weights
+  W1, b1, W2, b2 = model['W1'], model['b1'], model['W2'], model['b2']
+  N, C, H, W = X.shape
+
+  # We assume that the convolution is "same", so that the data has the same
+  # height and width after performing the convolution. We can then use the
+  # size of the filter to figure out the padding.
+  conv_filter_height, conv_filter_width = W1.shape[2:]
+  assert conv_filter_height == conv_filter_width, 'Conv filter must be square'
+  assert conv_filter_height % 2 == 1, 'Conv filter height must be odd'
+  assert conv_filter_width % 2 == 1, 'Conv filter width must be odd'
+  conv_param = {'stride': 1, 'pad': (conv_filter_height - 1) / 2}
+  pool_param = {'pool_height': 2, 'pool_width': 2, 'stride': 2}
+  dropout_param = {'p': dropout}
+  dropout_param['mode'] = 'test' if y is None else 'train'
+
[…]

### 2015-03-15 code/cs231n/convnet_wrapper.py

@@ -1,0 +1,306 @@
+"""
+from
+https://piazza.com/class/i37qi08h43qfv?cid=452
+
+ex:
+
+# define the layers, layers are defined from input to output
+layer_defs = [
+    {
+        'name': 'C0',
+        'type': 'conv-relu-pool',
+        'size': 3,
+        'channels': int(sys.argv[1]),
+    },
+
+    {
+        'name': 'C1',
+        'type': 'conv-relu-pool',
+        'size': 3,
+        'channels': int(sys.argv[2]),
+    },
+
+    {
+        'name': 'C2',
+        'type': 'conv-relu',
+        'size': 3,
+        'channels': int(sys.argv[3]),
+    },
+
+    {
+        'name': 'A0',
+        'type': 'affine-relu',
+        'outsize': int(sys.argv[4]),
+    },
+
+    {
+        'name': 'A1',
+        'type': 'affine',
+        'outsize': num_classes,
+    },
+]
+
+# initialize the network model and layers
+model, layers = init_modular_convnet(layer_defs)
+loss_func = modular_loss_function(layers)
+
+# train my awesome neural network here using loss_func for training
+# ...
+# ...
+
+# save the network to disk
+save_modular_convnet(best_model, layers, 'layer_model.pickle')
+"""
+
+import cPickle as pickle
+
+def raw_forward_function(layer_type):
+    """
+    Given a layer type as a string, return the corresponding forward function
+    """
+    forward_functions_dict = {
+        'conv': conv_forward_fast,
+        'relu': relu_forward,
+        'pool': max_pool_forward_fast,
+        'affine': affine_forward,
+        'affine-relu': affine_relu_forward,
+        'conv-relu': conv_relu_forward,
+        'conv-relu-pool': conv_relu_pool_forward,
+    }
+
+    return forward_functions_dict[layer_type]
+
+def raw_backward_function(layer_type):
+    """
+    Given a layer type as a string, return the corresponding backward function
+    """
+    backward_functions_dict = {
+        'conv': conv_backward_fast,
+        'relu': relu_backward,
+        'pool': max_pool_backward_fast,
+        'affine': affine_backward,
+        'affine-relu': affine_relu_backward,
+        'conv-relu': conv_relu_backward,
+        'conv-relu-pool': conv_relu_pool_backward,
+    }
+
+    return backward_functions_dict[layer_type]
+
+def extra_params(layer_type, layer_name, model):
+    """
[…]

### 2015-03-15 code/cs231n/data_augmentation.py

@@ -1,0 +1,171 @@
+import numpy as np
+
+
+def random_flips(X):
+  """
+  Take random x-y flips of images.
+
+  Input:
+  - X: (N, C, H, W) array of image data.
+
+  Output:
+  - An array of the same shape as X, containing a copy of the data in X,
+    but with half the examples flipped along the horizontal direction.
+  """
+  out = None
+  #############################################################################
+  # TODO: Implement the random_flips function. Store the result in out.       #
+  #############################################################################
+  p = 0.5
+  out = X.copy()
+  N = X.shape[0]
+  for n in xrange(N):
+      if np.random.rand() < p:
+          out[n] = out[n][:, :, ::-1]
+  #############################################################################
+  #                           END OF YOUR CODE                                #
+  #############################################################################
+  return out
+
+
+def random_crops(X, crop_shape):
+  """
+  Take random crops of images. For each input image we will generate a random
+  crop of that image of the specified size.
+
+  Input:
+  - X: (N, C, H, W) array of image data
+  - crop_shape: Tuple (HH, WW) to which each image will be cropped.
+
+  Output:
+  - Array of shape (N, C, HH, WW)
+  """
+  N, C, H, W = X.shape
+  HH, WW = crop_shape
+  assert HH < H and WW < W
+
+  out = np.zeros((N, C, HH, WW), dtype=X.dtype)
+  #############################################################################
+  # TODO: Implement the random_crops function. Store the result in out.       #
+  #############################################################################
+  for n in xrange(N):
+    X1 = X[n]
+    # crop is in 2nd and 3rd dims
+    # and can start anywhere from 0 to H-HH-1
+    HHn = np.random.randint(0, H - HH)  # high exclusive
+    WWn = np.random.randint(0, W - WW)  # high exclusive
+    out[n] = X[n, :, HHn:HH + HHn, WWn: WW + WWn]
+
+  #############################################################################
+  #                           END OF YOUR CODE                                #
+  #############################################################################
+
+  return out
+
+
+def random_contrast(X, scale=(0.8, 1.2)):
+  """
+  Randomly adjust the contrast of images. For each input image, choose a
+  number uniformly at random from the range given by the scale parameter,
+  and multiply each pixel of the image by that number.
+
+  Inputs:
+  - X: (N, C, H, W) array of image data
[…]

### 2015-03-15 code/cs231n/layers.py

@@ -1,0 +1,204 @@
+import numpy as np
+
+def affine_forward(x, w, b):
+  """
+  Computes the forward pass for an affine (fully-connected) layer.
+
+  The input x has shape (N, d_1, ..., d_k) where x[i] is the ith input.
+  We multiply this against a weight matrix of shape (D, M) where
+  D = \prod_i d_i
+
+  Inputs:
+  x - Input data, of shape (N, d_1, ..., d_k)
+  w - Weights, of shape (D, M)
+  b - Biases, of shape (M,)
+  
+  Returns a tuple of:
+  - out: output, of shape (N, M)
+  - cache: (x, w, b)
+  """
+  out = x.reshape(x.shape[0], -1).dot(w) + b
+  cache = (x, w, b)
+  return out, cache
+
+
+def affine_backward(dout, cache):
+  """
+  Computes the backward pass for an affine layer.
+
+  Inputs:
+  - dout: Upstream derivative, of shape (N, M)
+  - cache: Tuple of:
+    - x: Input data, of shape (N, d_1, ... d_k)
+    - w: Weights, of shape (D, M)
+
+  Returns a tuple of:
+  - dx: Gradient with respect to x, of shape (N, d1, ..., d_k)
+  - dw: Gradient with respect to w, of shape (D, M)
+  - db: Gradient with respect to b, of shape (M,)
+  """
+  x, w, b = cache
+  dx = dout.dot(w.T).reshape(x.shape)
+  dw = x.reshape(x.shape[0], -1).T.dot(dout)
+  db = np.sum(dout, axis=0)
+  return dx, dw, db
+
+
+def relu_forward(x):
+  """
+  Computes the forward pass for a layer of rectified linear units (ReLUs).
+
+  Input:
+  - x: Inputs, of any shape
+
+  Returns a tuple of:
+  - out: Output, of the same shape as x
+  - cache: x
+  """
+  out = np.maximum(0, x)
+  cache = x
+  return out, cache
+
+
+def relu_backward(dout, cache):
+  """
+  Computes the backward pass for a layer of rectified linear units (ReLUs).
+
+  Input:
+  - dout: Upstream derivatives, of any shape
+  - cache: Input x, of same shape as dout
+
+  Returns:
+  - dx: Gradient with respect to x
+  """
+  x = cache
+  dx = np.where(x > 0, dout, 0)
+  return dx
+
+
+def dropout_forward(x, dropout_param):
+  """
[…]

### 2015-03-15 doc/Course_Paper/paper.tex
updating paper

@@
     % TODO: Insert results!
+    We find that using a convolution neural network trained on Galaxy
+    morphologies as a feature extractor performs admirably but markedly worse
+    than the citizen scientists.
     Scripts used in the analysis of this paper are freely available at
@@
 
-A consequence of Einstein's Theory of General Relativity is that mass bends the
-path of light. Most of the time the deflections are very small; the original
-'gravitational lens' that tested the veracity of Einstein's theory in the years
-after the first world war was the sun, which deflected light from stars behind
-it only a few seconds of arc. However, when light passes through a particularly
-deep gravitational potential (say, near the center of the dark matter halo of a
-galaxy cluster), the deflections can be particularly large, resulting in
-brilliant arcs and multiple images. These strong deflections due to light
-passing through a deep gravitational potential are termed strong gravitational
-lenses. The gravitational lensing signal can heuristically be thought of as a
-trade off between a couple factors: larger gradients in the gravitational
-potential create larger distortions (and larger gradients in the gravitational
-potential tend to reside near to the center of the foreground galaxy or galaxy
-cluster); the more separated the foreground and background objects are, the
-bigger the proportional distortion on the background object by the foreground
-(but the farther away the background object is, the smaller it
-appears\footnote{Note that this is only true in the ``low redshift'' universe:
-  when objects are farther away than a cosmological redshift of $z \approx 2.5$
-or about 2.5 Gyr after the birth of the Universe, they will actually
-\textit{grow} in angular extent.}).
+A consequence of Einstein's Theory of General Relativity
+is that mass bends the path of light.\cite{Einstein:2007aa} Most of the time the deflections are very
+small; the original 'gravitational lens' that tested the veracity of Einstein's
+theory in the years after the first world war was the sun, which deflected
+light from stars behind it only a few seconds of arc. However, when light
+passes through a particularly deep gravitational potential (say, near the
+center of the dark matter halo of a galaxy cluster), the deflections can be
+particularly large, resulting in brilliant arcs and multiple images. These
+strong deflections due to light passing through a deep gravitational potential
+are termed strong gravitational lenses. The gravitational lensing signal can
+heuristically be thought of as a trade off between a couple factors: larger
+gradients in the gravitational potential create larger distortions (and larger
+gradients in the gravitational potential tend to reside near to the center of
+the foreground galaxy or galaxy cluster); the more separated the foreground and
+background objects are, the bigger the proportional distortion on the
+background object by the foreground (but the farther away the background object
+is, the smaller it appears\footnote{Note that this is only true in the ``low
+  redshift'' universe: when objects are farther away than a cosmological
+  redshift of $z \approx 2.5$ or about 2.5 Gyr after the birth of the Universe,
+they will actually \textit{grow} in angular extent.}).\cite{Kochanek:2006oz}
 
@@
 electromagnetic radiation but does have mass and hence interacts
-gravitationally with normal baryonic matter. This allows us to tally the mass
+gravitationally with normal baryonic matter.\cite{Refsdal:1964aa} This allows us to tally the mass
 of the largest gravitationally-bound structures in the universe, galaxy
@@
 the most massive objects in the universe can be. The properties of the bent
-light itself can also say much about that expansion history. When an object is
+light itself can also say much about that expansion
+history.\cite{Refsdal:1964ab, Kundic:1997aa, Barnacka:2014aa} When an object is
 strongly-lensed into multiple images, each image travels a different span of
@@
 poor completeness or poor purity -- and often poor both. Additionally, some
-algorithms are better at finding some types of lenses than others; some
-perform well on the brilliant arcs, but poorly on the multiply-imaged objects,
-or vice versa. New algorithms need to be developed to find more strong
-gravitational lenses, and more strong lenses need to be found to power these
-algorithms.
+algorithms are better at finding some types of lenses than others; some perform
+well on the brilliant arcs, but poorly on the multiply-imaged objects, or vice
+versa. For example, \cite{Kubo:2008aa} attempt to fit arc-like features in
+images in order to find strong gravitational lenses, but this means that
+multiply-imaged quasars are completely ignored. \cite{Agnello:2015aa} and
+\cite{Chan:2014aa} in contrast develop an algorithm for finding gravitationally
+lensed quasars based on catalog-level colors and magnitudes, precluding their
+algorithm finding strong gravitational lens arcs. New algorithms need to be
+developed to find more strong gravitational lenses, and more strong lenses need
+to be found to power these algorithms. These algorithms need to not only
+identify potential lenses accurately, but be able to make strong statements
+about their contamination rates, as spectroscopic follow-up can be an expensive
+endeavor. \cite{Momcheva:2015aa} performed spectroscopic follow-up on 9768
+galaxies, finding 28 new strong gravitational lens systems, but taking 40
+nights of telescope time on expensive telescopes.
 
@@
 that the $x$-axis stops at a false positive rate of 0.5, and the $y$-axis
-begins at a true positive rate of 0.5.}
+begins at a true positive rate of 0.5. Truly random guessing (which results in
+a 1:1 relationship between the true positive rate and the false positive rate)
+would not show up on this graph.}
 \label{fig:roc_tl} 
@@
 
-% TODO: Redescribe the network, including data augmentation and such.
-Transfer Learning provides a mechanism by which
+Transfer Learning relies on the idea that convolution neural networks that
+perform similar tasks pick up similar features, as well as the observation that
+lower levels in neural networks tend to be quite generic in the features they
+pick out. They provide an answer to the scenario when there are too few data to
+effectively train a complex system like deep convolution neural networks: start
+the training of your new system from the results of training a similar system.
+For the scope of this project, we chose to examine the effects of transfer
+learning from galaxy morphology to strong gravitational lens identification. In
+both cases, an input image of a galaxy is fed into the network, and some
+classification is read out. Additionally, both networks need to differentiate
+shapes in the central regions of the galaxy image (for example to find bars in
+spiral galaxies) from outer regions (spiral arms, arcs, multiply-lensed
+systems). The Galaxy Zoo competition provides an ideal candidate for transfer
+learning because of these facts, and also because the image quality is
+comparable between the Sloan Digital Sky Survey (the telescope survey on which
+the Galaxy Zoo images were based) and the CFHTLS survey. We have on hand a
+convolution neural network trained to classify galaxy morphology provided by
+Ryan Keisler \footnote{\url{rkeisler@stanford.edu}}. The architecture of that
+network can be seen in Figure~\ref{fig:ryan_architecture}. We do the simplest
+thing possible: we run the convolution neural network as a feature extractor,
+and take images from the fully connected layer. Thus we transform a
+$96\times96\times3$ image into a 500 feature vector. We then train these
+feature vectors on various classifiers (Random Forest, Support Vector Machine,
+Softmax) and evaluate results against a test set. We distinguish between stage
+1 and stage 2 data because the simulated lenses changed between the two sets.
+The code that produces the feature vectors also augments the data by
+automatically producing feature vectors of flips and rotations of the input
+images. This allows us to increase the size of our input dataset nearly
+20-fold.
+
+We train these datasets on three classifiers: Random Forests (which are an
+ensemble of decision trees trained on the data), Softmax and linear Support
+Vector Machines. We use stochastic gradient descent for the latter two
+classifiers. We create a test dataset by randomly extracting 20 percent of the
+dataset and setting it aside. We also ensure that any data augmentation stays
+in the training or test sets. Our goal with all the above systems is not to
+find the maximal accuracy, but to find some reasonable trade-off between the
+true positive rate and the false positive rate: we want to find as many lenses
+as we can, but we also know that confirmation of these lenses by spectroscopic
+follow-up is an expensive endeavor such that we want to minimize the number of
+non-lenses that make it into our candidate list. Because of this, any potential
+candidate list we would make from any of our classifiers has a relatively hard
+threshold at a false positive rate of 0.2.
+
+The resultant Receiver Operating Curves can be observed in
+Figure~\ref{fig:roc_tl}. In general we find that support vector machines
+perform the best as a classifier on the feature vectors, but that the feature
+vectors perform more poorly than the \textsc{Space Warps} users. We must caveat
+though that \textsc{Space Warps} does not create a validation dataset against
+which to test the performance of the system. Even if we don't perform quite as
+well as the citizen scientists, we consider this a
+promising baseline for future performance by transfer learning: we have not
+even begun to consider potential performance gains by retraining the
+convolution neural network on the \textsc{Space Warps} data.
 
@@
 
-% TODO: Remake with linear classifiers.
-% TODO: Include results from CNN as well.
-% TODO: Touch up description
-\begin{figure}%[!ht]
-\centering
-\epsfig{file=Figures/roc_curve.pdf,width=\linewidth,angle=0,clip=}
-\caption{Receiver Operating Curves for the Random
-  Forest Transfer Learning and the \textsc{Space Warps} systems, divided by
-  stage.}
-  \label{fig:roc}
-\end{figure}
-
-% TODO: make figure with probability distributions between classifiers and sw
+% Do we need this section?
 
@@
 % Discuss how well things went. Or didn't go.
+The need for new automated detection algorithms for finding strong
+gravitational lenses will only become more pressing in the next decade, as it
+becomes infeasible for scientists to scan images by eye for such systems. Using
+the \textsc{Space Warps} dataset, we have examined how convolution neural
+networks trained both on this particular dataset and on other datasets can
+perform at the detection task.
+% TODO: brief blurb about our CNN
+We find that features extracted from a convolution neural network trained on
+the classification of galaxy morphology (with a linear support vector
+classifier on top for converting the feature vector to a binary ``lens'' and
+``not-lens'') performs admirably but markedly worse than the citizen scientists
+trained on the dataset. Further work examining improvements by retraining the
+whole neural network could lead to a generic classification machine that takes
+images from \textit{any} galaxy survey and states whether the image contains a
+strong lens or not. Additionally there is much potential in both direct
+convolution neural networks and transfer learning from other networks in
+linking the classification outputs of the networks with the probability
+estimates of the \textsc{Space Warps} system, which also examined several
+thousands more ``unknown'' systems and could lead to more gravitational lenses
+being identified.
 
@@
 %-------------------------------------------------------------------------
-\section{Intermediate/Preliminary Results}
-
-We have created the cutouts of our images from the catalogs.
-We have begun constructing our own convolutional network with two
-convolutional/max-pool layers and two fully-connected layers. Although we have
-not yet trained on our images, we can load these images and run them
-through our initialized classifier.
+% References
 

### 2015-03-15 doc/Course_Paper/paper.tex
more updates

@@
 \usepackage{amssymb}
-
+\input{macros.tex}
 % Include other packages here, before hyperref.
@@
     morphologies as a feature extractor performs admirably but markedly worse
-    than the citizen scientists.
+    than the citizen-scientists.
     Scripts used in the analysis of this paper are freely available at
@@
 
-A consequence of Einstein's Theory of General Relativity
-is that mass bends the path of light.\cite{Einstein:2007aa} Most of the time the deflections are very
+A consequence of Einstein's Theory of General Relativity is that mass bends the
+path of light. \cite{Einstein:2007aa} Most of the time the deflections are very
 small; the original 'gravitational lens' that tested the veracity of Einstein's
@@
 redshift of $z \approx 2.5$ or about 2.5 Gyr after the birth of the Universe,
-they will actually \textit{grow} in angular extent.}).\cite{Kochanek:2006oz}
+they will actually \textit{grow} in angular extent.}). \cite{Kochanek:2006oz}
 
@@
 electromagnetic radiation but does have mass and hence interacts
-gravitationally with normal baryonic matter.\cite{Refsdal:1964aa} This allows us to tally the mass
-of the largest gravitationally-bound structures in the universe, galaxy
-clusters, which can give us insight into the formation history of these massive
-objects. In this way, strong gravitational lenses can then tell us something
-about the expansion history of the universe, by setting limits on how massive
-the most massive objects in the universe can be. The properties of the bent
-light itself can also say much about that expansion
-history.\cite{Refsdal:1964ab, Kundic:1997aa, Barnacka:2014aa} When an object is
-strongly-lensed into multiple images, each image travels a different span of
-space and time. When an object does not vary much with time, these different
-path lengths have no practical import. However, if the object varies
-appreciably quickly (say it is a distant supermassive black hole at the center
-of a galaxy whose accretion disk emits high-energy radiation at varying rates)
-then these different path lengths can be used to pin down the rate of expansion
-of the universe.
+gravitationally with normal baryonic matter. \cite{Refsdal:1964aa} This allows
+us to tally the mass of the largest gravitationally-bound structures in the
+universe, galaxy clusters, which can give us insight into the formation history
+of these massive objects. \cite{Oguri:2010aa} They can also find extremely
+distant objects. \cite{Hezaveh:2014aa} In this way, strong gravitational lenses
+can then tell us something about the expansion history of the universe, by
+setting limits on how massive the most massive objects in the universe can be.
+The properties of the bent light itself can also say much about that expansion
+history. \cite{Refsdal:1964ab, Kundic:1997aa, Gonzalez:2012aa, Barnacka:2014aa,
+Hojjati:2014aa, Suyu:2014aa} When an object is strongly-lensed into multiple
+images, each image travels a different span of space and time. When an object
+does not vary much with time, these different path lengths have no practical
+import. However, if the object varies appreciably quickly (say it is a distant
+quasar, a supermassive black hole at the center of a galaxy whose accretion
+disk emits high-energy radiation at varying rates) then these different path
+lengths can be used to pin down the rate of expansion of the universe. Finally,
+strong gravitational lenses also have the potential to rule out or validate
+alternatives to General Relativity. \cite{Troxel:2014aa}
 
@@
 extremely rare. A next generation optical survey like the Large Synoptic Survey
-Telescope can expect to find only ten thousand lenses in the whole sky, while
-it will find ten billion galaxies. Currently in astronomy there are only order
-hundreds of strong gravitational lenses known, mostly discovered by 'eyeball
-squads' of graduate students. The small number means that target criteria must
-be somewhat broad in order to maintain a relatively high completeness. Using
-reasonable target criteria to find strong lenses such as looking only at
-massive galaxies still means that nearly ten million objects will need to be
-inspected in the next generation in order to find those ten thousand lenses. A
-team of ten graduate students could expect to spend about 14 years looking at
-these objects. Computer algorithms are not much better: most current machine
-learning algorithms are woefully-underpowered for this task, and generally have
-poor completeness or poor purity -- and often poor both. Additionally, some
+Telescope or the Euclid space telescope can expect to find only ten thousand
+lenses in the whole sky, while it will find ten billion galaxies.
+\cite{Serjeant:2014aa} Currently in astronomy there are only order hundreds of
+strong gravitational lenses known, mostly discovered by 'eyeball squads' of
+graduate students. The small number means that target criteria must be somewhat
+broad in order to maintain a relatively high completeness. Using reasonable
+target criteria to find strong lenses such as looking only at massive galaxies
+still means that nearly ten million objects will need to be inspected in the
+next generation in order to find those ten thousand lenses. A team of ten
+graduate students could expect to spend about 14 years looking at these
+objects. Computer algorithms are not much better: most current machine learning
+algorithms are woefully-underpowered for this task, and generally have poor
+completeness or poor purity -- and often poor both. Additionally, some
 algorithms are better at finding some types of lenses than others; some perform
@@
 automated detection of strong lens systems. We will also assess the performance
-of citizen scientists by comparing our results to them. From other graduate
+of citizen-scientists by comparing our results to them. From other graduate
 work (but not coursework), we have the locations and categories of around one
@@
 constructed approximately 30000 cutouts sized $96\times96\times3$. These cutouts
-are selected based on where citizen scientists clicked, on the theory that both
+are selected based on where citizen-scientists clicked, on the theory that both
 'correct' and 'incorrect' selections provide useful information about the
@@
 a second 'refinement stage' of the project, where much harder simulations were
-given to users. The 'test' images are the fields that citizen scientists
+given to users. The 'test' images are the fields that citizen-scientists
 viewed, assessing whether a lens was in the field or not. In these 'test'
@@
 which to test the performance of the system. Even if we don't perform quite as
-well as the citizen scientists, we consider this a
+well as the citizen-scientists, we consider this a
 promising baseline for future performance by transfer learning: we have not
@@
 
-% Do we need this section?
+We briefly attempt to answer the questions we posed at the beginning of this
+study:
+
+% answer the questions
+\subsection*{Do we have enough data to reasonably train and test a CNN?}
+
+We do, and we have easy ways to augment the data through rotations, flips, and
+(in the future) other transformations that leave the identification of a strong
+gravitational lens invariant.
+
+\subsection*{How do citizen-scientists do compared with this automated system?}
+
+Currently the citizen-scientists outperform our automated systems.
+
+\subsection*{Can we use the results of citizen-scientists to train the CNN?}
+
+In the sense that the citizen-scientists are used in creating the cutouts for
+our system, we find their results quite useful. We were unable to use them for
+much more than that, however. Future work could involve calibrating our
+classifications against the probabilities from the \textsc{Space Warps}
+citizen-scientist classification system.
+
+\subsection*{How well does using features extracted from a convolution neural
+network trained on galaxy morphology perform when determining the presence of
+strong gravitational lenses?}
+
+We appear to do admirably well in that our receiver operating curves obtain
+remarkably high true positive rates for modestly low false positive rates.
+However, the extracted features do worse than the citizen-scientists from
+\textsc{Space Warps}.
 
@@
 classifier on top for converting the feature vector to a binary ``lens'' and
-``not-lens'') performs admirably but markedly worse than the citizen scientists
+``not-lens'') performs admirably but markedly worse than the citizen-scientists
 trained on the dataset. Further work examining improvements by retraining the
@@
 %-------------------------------------------------------------------------
-\section{Acknowledgements}
+\subsection*{Acknowledgements}
 We received access to the dataset through Phil Marshall

### 2015-03-16 doc/Course_Paper/paper.tex
more updates to figure

@@
   morphologies. We find that of the linear classifiers on the feature vectors,
-support vector machine classifications perform best on the test dataset,
+Softmax classifications perform best on the test dataset,
 however all the feature vectors perform worse than the users themselves. Note

### 2015-03-16 code/analyze_TL_database.py
paper draft

@@ -208,5 +208,11 @@ for stage in [1, 2]:
 # plt.show()
 
 
+bindir = project_dir + 'binarys/'
+np.save(bindir + 'fpr', fpr)
+np.save(bindir + 'tpr', tpr)
+np.save(bindir + 'cuts', cuts)
+np.save(bindir + 'scores', scores)
+np.save(bindir + 'thresh', thresh)
 
 import ipdb; ipdb.set_trace()

### 2015-03-16 doc/Course_Paper/paper.tex
more paper

@@
 are termed strong gravitational lenses. The gravitational lensing signal can
-heuristically be thought of as a trade off between a couple factors: larger
+heuristically be thought of as a trade off between a couple of factors: larger
 gradients in the gravitational potential create larger distortions (and larger
 gradients in the gravitational potential tend to reside near to the center of
-the foreground galaxy or galaxy cluster); the more separated the foreground and
+the foreground galaxy or galaxy cluster), while the more separated the foreground and
 background objects are, the bigger the proportional distortion on the
-background object by the foreground (but the farther away the background object
+background object by the foreground (conversely, the farther away the background object
 is, the smaller it appears\footnote{Note that this is only true in the ``low
@@
 From approximately 12000 fields of $440\times440\times3$ fields, we have
-constructed approximately 30000 cutouts sized $96\times96\times3$. These cutouts
-are selected based on where citizen-scientists clicked, on the theory that both
-'correct' and 'incorrect' selections provide useful information about the
-characteristics of gravitational lenses. In general, we have access to two
-broad classes of images: 'training' and 'test' images. The 'training' images
+constructed approximately 30000 cutouts sized $96\times96\times3$. These
+cutouts are selected based on where citizen-scientists clicked, on the theory
+that both `correct' and `incorrect' selections provide useful information about
+the characteristics of gravitational lenses. In general, we have access to two
+broad classes of images: `training' and `test' images. The `training' images
 include fields that were verified in advance to not contain any lenses as well
@@
 objects are over-exaggerated and extremely obvious, but we also have access to
-a second 'refinement stage' of the project, where much harder simulations were
-given to users. The 'test' images are the fields that citizen-scientists
+a second `refinement stage' of the project, where much harder simulations were
+given to users. The `test' images are the fields that citizen-scientists
 viewed, assessing whether a lens was in the field or not. In these 'test'
@@
 
-% TODO: Insert figure for architecture
+\begin{figure}%[!ht]
+\centering
+\epsfig{file=Figures/theano_roc.pdf,width=\linewidth,angle=0,clip=}
+\caption{Receiver Operating Curves extracted from the training and validation sets used with 
+our Convolutional Neural Network after training for 35 epochs. These results should be compared 
+with the \textsc{Space Warps} Stage 1 Receiver Operating Curve in Figure~\ref{fig:roc_tl}, which 
+outperforms our network for false positive rates below $\sim 0.3$. However, above this 
+point our Convolutional Net achieves higher true positive rates than any of the other (machine 
+or human) methods herein considered.}
+\label{fig:roc_theano}
+\end{figure}
+
+In order to retain our ability to experiment with non-traditional training methods and 
+architectures, we chose to implement our Convolutional Neural Network (CNN) from the ground 
+up using \textsc{theano}. We settled on a five-layer layer architecture consisting of two 
+convolution/max dropout layers, a fully connected layer, and a softmax layer. Training was 
+carried out using RMSprop with a decay rate of 0.9, and the model was regularized using $L_2$ normalization 
+with a regularization strength of $10^{-4}$ and dropout with probability 0.5 at each layer . We chose our convolution layers to have stride 4 and 
+depth 10 (with padding to maintain the input image size), and the subsequent max pooling layers to have stride 2. Assigning our fully 
+connected layer 50 neurons (and including the paramters in the subsequent softmax layer) gave 
+our model a total of 288,492 trainable parameters.
+
+Due to memory and time constraints, we have not yet trained this CNN on all our of data. The results 
+of training it on 8,000 randomly selected images from the \textsc{Space Warps} Stage 1 data set for 
+35 epochs is shown in Figure~\ref{fig:roc_theano}, where we have plotted the Receiver Operating Curve 
+obtained from our model's predictions on the training set and a test set composed of another (non-overlapping)
+ randomly selected 8,000 images from the \textsc{Space Warps} Stage 1 data set. This result should be easily improved on by training on more of 
+ our images, by adding data augmentation, and by increasing the complexity of our CNN architecture. However, it already shows that training 
+ a CNN solely on the \textsc{Space Warps} data set is a viable option, given enough CPU power and time.
 
@@
 
-% TODO: Touch up description
 \begin{figure}%[!ht]
@@
 \subsection*{Acknowledgements}
-We received access to the dataset through Phil Marshall
-\footnote{\url{pjm@slac.stanford.edu}}, who also graciously explained to us how
+We received access to the dataset through Phil Marshall\footnote{\url{pjm@slac.stanford.edu}}, who also graciously explained to us how
 \textsc{Space Warps}
 currently works. The convolutional neural network upon which the transfer
-learning is based was kindly provided to us by Ryan Keisler
-\footnote{\url{rkeisler@stanford.edu}}.
+learning is based was kindly provided to us by Ryan Keisler\footnote{\url{rkeisler@stanford.edu}}.
 All errors in interpretation or otherwise are entirely our own.

### 2015-03-16 doc/Course_Paper/paper.tex
updating code. added one more reference. pushing final paper tex and pdf

@@
 perform at the detection task.
-% TODO: brief blurb about our CNN
-We find that features extracted from a convolution neural network trained on
+
+While our Convolutional Neural Net trained on just the \textsc{Space Warps} dataset 
+has not yet outperformed citizen-scientists at sufficiently low false positive rates, there is 
+ample reason to believe marked improvements can be made with more time and 
+CPU power. In addition to training on more images (and augmented images) for longer, 
+we can expect to gain a significant reduction in training time by taking advantage of 
+\textsc{theano}'s GPU capabilities and by implementing batch normalization. \cite{Ioffe:2015aa} 
+It may also be beneficial to explore larger and more expressive CNN architectures.
+
+In our application of Transfer Learning, we find that features extracted from a convolution neural network trained on
 the classification of galaxy morphology (with a linear support vector
@@
 
-% TODO: copy relevant bib stuff over so I don't have a 4 mb bib
 {\small
